
\documentclass[nooutcomes]{ximera}
%\documentclass[space,handout,nooutcomes]{ximera}

\input{../preamble.tex}

\title{Numbers}
\author{Bart Snapp and Brad Findell and Jenny Sheldon}
\begin{document}
\begin{abstract}
Numbers you should know. 
\end{abstract}
\maketitle


%\begin{problem}
%Problem
%\begin{freeResponse}
%\begin{hint}
%Hint
%\end{hint}
%\end{freeResponse}
%\end{problem} 


We assume you know your multiplication facts up to $12\times 12$. Because you are math teachers, it helps to have some other numbers on the tip of your tongue---or at least be able to figure them out quickly in your head. 


\begin{problem}
Cubes:
\begin{enumerate}
\item $2^3 = \answer{8}$
\item $3^3 = \answer{27}$
\item $4^3 = \answer{64}$
\item $5^3 = \answer{125}$
\item $6^3 = \answer{216}$
\item $7^3 = \answer{343}$
\item $8^3 = \answer{512}$
\item $9^3 = \answer{729}$
\item $10^3 = \answer{1000}$
\item $11^3 = \answer{1331}$
\end{enumerate}
\end{problem}

\begin{problem}
Factorials: $4!$ means $4\cdot3\cdot2\cdot1$.  
\begin{enumerate}
\item $4!= \answer{24}$
\item $5!= \answer{120}$
\item $6!= \answer{720}$
\item $7!= \answer{5040}$
\item $3!= \answer{6}$
\item $2!= \answer{2}$
\item $1!= \answer{1}$
\item $0!= \answer{1}$
\end{enumerate}
\end{problem}

\begin{problem}
Powers of 2
\begin{enumerate}
\item $2^4 = \answer{16}$
\item $2^5 = \answer{32}$
\item $2^6 = \answer{64}$
\item $2^7 = \answer{128}$
\item $2^8 = \answer{256}$
\item $2^9 = \answer{512}$
\item $2^{10} = \answer{1024}$
\end{enumerate}
\end{problem}

\begin{problem}
Because computers are designed around base $\answer[format=string]{two}$, the powers of two turn out to be very important.  In particular, because $2^{10}=\answer{1024}$, which is approximately $1000$, computer science borrowed the base-ten prefixes from the metric system.  

Computer memory and storage is typically measured in bytes, where 1 byte can store one ``character,'' such as a letter, punctuation mark, or space in this sentence.  With the metric prefixes, we approximate powers of 2 with powers of ten.  For example: 
\begin{enumerate}
\item $1$ Kilobyte (1 KB) $=  2^{10}\textrm{ bytes }\approx \answer{10^3}$ bytes = one \wordChoice{\choice[correct]{thousand} \choice{million} \choice{billion} \choice{trillion}} bytes.
\item $1$ Megabyte (1 MB) $=  \answer{2^{20}} \textrm{ bytes } \approx \answer{10^6}$ bytes = one \wordChoice{\choice{thousand} \choice[correct]{million} \choice{billion} \choice{trillion}} bytes.
\item $1$ Gigabyte (1 GB) $=  \answer{2^{30}} \textrm{ bytes } \approx \answer{10^9}$ bytes = one \wordChoice{\choice{thousand} \choice{million} \choice[correct]{billion} \choice{trillion}} bytes.
\item $1$ Terabyte (1 TB) $=  \answer{2^{40}} \textrm{ bytes } \approx \answer{10^{12}}$ bytes = one \wordChoice{\choice{thousand} \choice{million} \choice{billion} \choice[correct]{trillion}} bytes.
\end{enumerate}

And 1 byte is composed of 8 bits, each of which is a base-two digit, either 0 or 1.  Today, internet speeds are usually given in Mb/sec, which means ``Megabits per second.'' So if your Internet speed is $100$ Mb/sec, 
that means $\answer{100\cdot10^6}$ bits per second.  (Type your answer without commas.)

Computer processor speeds are typically given in GHz, where 1 Hz = 1 cycle per second.  Many processors operations require 1 cycle to complete, so we imagine that a 2.4 GHz processor completes $\answer{2.4\cdot10^9}$ operations per second.  

Most modern processors have several `cores' which can perform operations at the same time.  But never mind.  

In 1984, a fully-equipped IBM PC/XT had $640$ KB of memory, a $10$ MB hard drive, and processor running at $4.77$ MHz.  In those years, it was possible to connect others through an ``online service provider'' (the World Wide Web did not exist), with a modem operating at 300 bits per second, tying up the phone line, and at a cost of $\$6$---or even as much as $\$30$ per hour.  (A leading online service provider was Compuserve, based in Columbus, Ohio.)  

In contrast, a typical 2021 laptop computer might have a 2.4 GHz processor, 16 GB of memory, a 512 GB (solid state) ``hard'' drive, and an internet connection operating at $100$ Mb/sec.  

By what factors do the capacities of a 2021 computer improve on the 1984 version?  

\begin{enumerate}
\item Processor speed: $\answer[tolerance=20]{2.4\cdot10^9/(4.77\cdot10^6)}$.  % 503.144654088050314
\item Disk storage: $\answer[tolerance=2000]{512\cdot10^9/(10\cdot10^6)}$.     % 51200.
\item Memory: $\answer[tolerance=1000]{16\cdot10^9/(640\cdot10^3)}$.           % 25000.
\item Online speed: $\answer[tolerance=17000]{100\cdot10^6/300}$.              % 333333.333333
\end{enumerate}
\end{problem}

\end{document}



