
\documentclass[nooutcomes]{ximera}
%\documentclass[space,handout,nooutcomes]{ximera}

\input{../preamble.tex}

\title{Home Base, Part A}
\author{Bart Snapp and Brad Findell}
\begin{document}
\begin{abstract}
Beginning problems about numbers in various bases. 
\end{abstract}
\maketitle

%For practice counting in bases other than ten, go to http://tube.geogebra.org/m/1529377.
%https://www.geogebra.org/m/Cz5myDUU
%
%For practice converting between bases, go to http://tube.geogebra.org/m/1527705.
%https://www.geogebra.org/m/ZEdZjfxm
%

If you haven't already practiced, take an opportunity now.  

\begin{center}
\geogebra{1529377}{995}{370}
\end{center}

\begin{center}
\geogebra{1527705}{780}{240}
\end{center}

Note:  The ``free response'' answers are not checked for accuracy.  To optimize your learning, submit your own answer before revealing the hint.  

\begin{problem}
To optimize my learning, I plan to $\answer[format=string]{submit}$ my own answer \wordChoice{\choice[correct]{before}\choice{after}} revealing the $\answer[format=string]{hint}$.  
\end{problem}

\begin{problem}Explain why the following ``joke'' is ``funny:'' \textit{There
  are $10$ types of people in the world. Those who understand base two
  and those who don't.}
\begin{freeResponse}
\begin{hint}
In base two, 10 is actually two.  So people who do not understand base two will not get the joke.  
\end{hint}
\end{freeResponse}
\end{problem} 

\begin{problem}You meet some Tripod aliens, they tally by threes. Thankfully
  for everyone involved, they use the symbols $0$, $1$, and $2$. 
\begin{enumerate}
%\item Can you explain how a Tripod would count from $11$ to $201$? Be
%  sure to carefully explain what number comes after $22$.
  \item Demonstrate how a Tripod would count, beginning at $11$.  
  
$11, \answer{12}, \answer{20}, \answer{21}, \answer{22}, \answer{100}, \answer{101}, 
\answer{102}, \\
\answer{110}, \answer{111}, \answer{112}, \answer{120}, \answer{121}, \answer{122}, \answer{200}, \answer{201}$

\item What number comes immediately before $10$?  $\answer{2}$
\item Before $210$? $\answer{202}$
\item Before $20110$? $\answer{20102}$
  Explain your reasoning.
\end{enumerate}
\end{problem} 

\begin{problem}You meet some people who tally by sevens. They use the symbols
  $O$, $A$, $B$, $C$, $D$, $E$, and $F$, in that order. (Note: Although it is common to use the letters $A$ through $F$ for digits greater than ten, these people are doing something different.)
\begin{enumerate}
\item What do the individual symbols $O$, $A$, $B$, $C$, $D$, $E$, and
  $F$ mean?  (Note $O$ is not $0$.)
  \begin{freeResponse}
    \begin{hint}
      0, 1, 2, 3, 4, 5, and 6, respectively.
    \end{hint}
  \end{freeResponse}
\item Demonstrate how to count from $DD$ to $AOC$?  (Note: Case matters.)

$DD, \answer[format=string]{DE}, \answer[format=string]{DF}, \answer[format=string]{EO}, \answer[format=string]{EA}, \\
\answer[format=string]{EB}, \answer[format=string]{EC}, \answer[format=string]{ED}, \answer[format=string]{EF}, \answer[format=string]{FO}, \\
\answer[format=string]{FA}, \answer[format=string]{FB}, \answer[format=string]{FC}, \answer[format=string]{FD}, \answer[format=string]{FE}, \\
\answer[format=string]{FF}, \answer[format=string]{AOO}, \answer[format=string]{AOA}, \answer[format=string]{AOB}, \answer[format=string]{AOC}$
\item What number comes immediately before $AO$?  $\answer[format=string]{F}$
\item Before $ABO$?  $\answer[format=string]{AAF}$
\item Before $EOFFA$? $\answer[format=string]{EOFFO}$
\end{enumerate}
\end{problem} 

\begin{problem}Now, suppose that you meet a hermit who tallies by
  thirteens. Demonstrate the hermit's counting below.    
  
$8, 9, \answer[format=string]{A}, \answer[format=string]{B}, \answer[format=string]{C}, \answer[format=string]{10}, \answer[format=string]{11}, \answer[format=string]{12}, \dots, \\
18, \answer[format=string]{19}, \answer[format=string]{1A}, \answer[format=string]{1B}, \answer[format=string]{1C}, \answer[format=string]{20}, \answer[format=string]{21}$

\end{problem} 


\end{document}